---
title: "Data_Mining_Instruction"
author: "Xuan"
date: "November 28, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read Me

This is an appending document in my personal Overwatch Project to help you understand how I built my dataset. 

Please note that the ranking system is a live update system. This means that everytime you run my code, it is highly possible that you acquire a different dataset from mine. The datasets I applied are stored in the repo, for your reference.

The recording time of the ranking system in my dataset is 11/30/2018.

## Code Sample
```{r eval=FALSE, include=FALSE}
library(rvest)
library(tidyverse)

# web scrapping
htmlana <- "https://www.overbuff.com/heroes/ana/rankings"
anaplayerid <- htmlana%>%read_html()%>%html_nodes(".sortable a")%>%html_attr("href")%>%as.data.frame()
anaplayerid <-unique(anaplayerid)
rownames(anaplayerid) <- c(1:100)

#check the dataset, and we detect a problem that special ids do not show correctly. For simplicity, we remove these rows.
anaplayerid <-data.frame(anaplayerid[-c(18,48,75,99),])
colnames(anaplayerid) <- "ID"
rownames(anaplayerid) <- c(1:96)
anaplayerid[,1] <-as.character(anaplayerid[,1])

#create links to their main pages
linkhead <- "https://www.overbuff.com"
linktailana <- "/heroes/ana?mode=competitive"
anaplayerid$PLAYER <- NA
for(i in 1:96){
anaplayerid$PLAYER[i] <- unlist(strsplit(anaplayerid[i,1], split='/', fixed=TRUE))[4] #list player-id if you are interested in some specific players
anaplayerid[i,1]<-sprintf("%s%s%s",linkhead,anaplayerid[i,1],linktailana)
}
colnames(anaplayerid) <- c("URL","ID")

#Game Record
#create empty data frame
GameRecordS13Ana <- as.data.frame(matrix(ncol=22,nrow=1,NA))
colnames(GameRecordS13Ana)<- c("GAME_PLAYED","WIN_RATE","ON_FIRE","ELIS","OBJ_KILLS","OBJ_TIME","DAMAGE","HEALING","DEATH","WEAPON_ACC","SCOPED_ACC","ENMY_SLEPT","OFF_ASSISTS","DEF_ASSISTS","BOOST_ASSISTS","ED","VOTING_CARDS","MEDALS","GOLD","SILVER","BRONZE","PLAYER_ID")

for (i in 1:96){
  tryCatch({
url <-  anaplayerid[i,1]
templink <- read_html(url)
a <- templink%>%html_nodes(".player-heroes .value")%>%html_text()
a <- as.data.frame(matrix(ncol=21,a,byrow=TRUE))
colnames(a)<- c("GAME_PLAYED","WIN_RATE","ON_FIRE","ELIS","OBJ_KILLS","OBJ_TIME","DAMAGE","HEALING","DEATH","WEAPON_ACC","SCOPED_ACC","ENMY_SLEPT","OFF_ASSISTS","DEF_ASSISTS","BOOST_ASSISTS","ED","VOTING_CARDS","MEDALS","GOLD","SILVER","BRONZE")
a$PLAYER_ID <-anaplayerid[i,2]
GameRecordS13Ana <- rbind.data.frame(a,GameRecordS13Ana) #add game records each by each
},error = function(e) {}) #avoid error situations
}

#convert factors to numeric values
GameRecordS13Ana$WIN_RATE<- lapply(GameRecordS13Ana$WIN_RATE, function(x) as.numeric(sub("%", "", x))) 
GameRecordS13Ana$WIN_RATE <- as.numeric(GameRecordS13Ana$WIN_RATE)/100
GameRecordS13Ana <- GameRecordS13Ana %>%filter(WIN_RATE<1.1) #remove outliers
GameRecordS13Ana$GAME_PLAYED <- as.numeric(GameRecordS13Ana$GAME_PLAYED)
GameRecordS13Ana$WIN <- round(GameRecordS13Ana$GAME_PLAYED*GameRecordS13Ana$WIN_RATE,0)
GameRecordS13Ana$LOSS <-GameRecordS13Ana$GAME_PLAYED-GameRecordS13Ana$WIN
GameRecordS13Ana$ED <- as.character(GameRecordS13Ana$ED)
GameRecordS13Ana$ED <- as.numeric(GameRecordS13Ana$ED)
GameRecordS13Ana$MEDALS <- as.character(GameRecordS13Ana$MEDALS)
GameRecordS13Ana$MEDALS <- as.numeric(GameRecordS13Ana$MEDALS)

#write down the file
write.csv(GameRecordS13Ana,file="GameRecordS13Ana.csv")


##Now we do this again to get the dataset for Reinhardt mains
# web scrapping
htmlrein <- "https://www.overbuff.com/heroes/reinhardt/rankings"
reinplayerid <- htmlrein%>%read_html()%>%html_nodes(".sortable a")%>%html_attr("href")%>%as.data.frame()
reinplayerid <-unique(reinplayerid)
rownames(reinplayerid) <- c(1:100)
reinplayerid <-as.data.frame(reinplayerid[-51,])
colnames(reinplayerid) <- "ID"
reinplayerid$ID <-as.character(reinplayerid$ID)


#create links to their main pages
linkhead <- "https://www.overbuff.com"
linktailrein <- "/heroes/reinhardt?mode=competitive"
reinplayerid$PLAYER <- NA
for(i in 1:99){
reinplayerid$PLAYER[i] <- unlist(strsplit(reinplayerid[i,1], split='/', fixed=TRUE))[4] #list player-id if you are interested in some specific players
reinplayerid[i,1]<-sprintf("%s%s%s",linkhead,reinplayerid[i,1],linktailrein)
}
colnames(reinplayerid) <- c("URL","ID")

GameRecordS13R <- as.data.frame(matrix(ncol=21,nrow=1,NA))
colnames(GameRecordS13R)<- c("GAME_PLAYED","WIN_RATE","ON_FIRE","ELIS","OBJ_KILLS","OBJ_TIME","DAMAGE","DEATH","SOLO_KILLS","FINAL_BLOWS","BLOCKED","CHARGE","ULT","FIRE","ED","VOTING_CARDS","MEDALS","GOLD","SILVER","BRONZE","PLAYER_ID")


for (i in 1:96){
  tryCatch({
url <-  reinplayerid[i,1]
templink <- read_html(url)
a <- templink%>%html_nodes(".player-heroes .value")%>%html_text()
a <- as.data.frame(matrix(ncol=20,a,byrow=TRUE))
colnames(a)<- c("GAME_PLAYED","WIN_RATE","ON_FIRE","ELIS","OBJ_KILLS","OBJ_TIME","DAMAGE","DEATH","SOLO_KILLS","FINAL_BLOWS","BLOCKED","CHARGE","ULT","FIRE","ED","VOTING_CARDS","MEDALS","GOLD","SILVER","BRONZE")
a$PLAYER_ID <-reinplayerid[i,2]
GameRecordS13R <- rbind.data.frame(a,GameRecordS13R)
},error = function(e) {})
}

#convert factors to numeric values
GameRecordS13R$WIN_RATE<- lapply(GameRecordS13R$WIN_RATE, function(x) as.numeric(sub("%", "", x))) 
GameRecordS13R$WIN_RATE <- as.numeric(GameRecordS13R$WIN_RATE)/100
GameRecordS13R <- GameRecordS13R %>%filter(WIN_RATE<1.1)
GameRecordS13R$GAME_PLAYED <- as.character(GameRecordS13R$GAME_PLAYED)
GameRecordS13R$GAME_PLAYED <- as.numeric(GameRecordS13R$GAME_PLAYED)
GameRecordS13R$WIN <- round(GameRecordS13R$GAME_PLAYED*GameRecordS13R$WIN_RATE,0)
GameRecordS13R$LOSS <-GameRecordS13R$GAME_PLAYED-GameRecordS13R$WIN 
GameRecordS13R$ED <- as.character(GameRecordS13R$ED)
GameRecordS13R$ED <- as.numeric(GameRecordS13R$ED)
GameRecordS13R$MEDALS <- as.character(GameRecordS13R$MEDALS)
GameRecordS13R$MEDALS <- as.numeric(GameRecordS13R$MEDALS)

#write down the file
write.csv(GameRecordS13R,file="GameRecordS13R.csv")

```
